Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint

### 2014-11-14 egpaper_final.tex

%   In contrast to other mid-level patch based 3D matching methods which requires extensive training and more than 10Gb of memory space and physical storage, ours require no training nor saving the templates thus enabling it to run on any personal computer. 
\end{abstract}

@@

\textcolor{red}{Should we mention how our method improves on this (ie by using CG)?}
% Using HOG
% In this paper, we also used Histograms of Orientations feature \cite{Dalal05} but in general, we could use any dense features.

@@

\begin{table*}[!htbp]
    \begin{center}
\begin{tabular}{|c|c|r|c|r|}
\hline
Methods (AP/MPPE) & before calibration  & time & after Calibration \cite{Aubry14} & time \\
\hline\hline
HOG\cite{Dalal05}     & 72.3 / 65.0           &  31ms  & 60.4 / 50.2                 & 8.7 sec \\ 
WHO\cite{Hariharan12} & 82.1 / 85.4           &  3811ms& 84.4 / 83.0                 & 12.4 sec  \\
WHO-CG                & 81.7 / \textbf{84.9}  &  104ms & 83.7 / 87.3                 & 8.3 sec \\
WHO-CG-Z              & 54.4 / 65.1           &  103ms & \textbf{92.8} / 86.7        & 8.7 sec  \\
% NZC-WHO-Z    & 89.10/\textbf{78.64} &    & 91.15/74.79                  &     \\
NZ-WHO-CG             & \textbf{90.0} / 82.8  &   79ms & 90.3 / \textbf{86.8}        & 8.5 sec   \\
\hline
\end{tabular}
\end{center}
\caption{Average Precision(AP) and Mean Precision in Pose Estimation (MPPE) \cite{Lopez-Sastre11} variations of WHO on 3DObject Car dataset\cite{Savarese07}. WHO refers to standard WHO using the method presented in \cite{Hariharan12}, WHO-CG uses iterative Conjugate Gradient method to generate WHO. WHO-CG-Z uses whiten the whole template and zero out textureless region. NZ-WHO-CG is the NZ-WHO which whitens only non-zero cells using iterative Conjugate Gradient method. The time column indicates the time to generate one template. We followed calibration procedure presented in \cite{Aubry14}.}
\label{tab:who_initializations}
\end{table*}

@@

% Estimating viewpoint continuously to fine tune a initial matching 
Since creating a template and validating the template become computationally inexpensive, making a proposal and validating the proposal to fine tune a 2D-3D matching became affordable. We used Metropolis-Hastings algorithm to generate a proposal pose and focal length and validate the proposal using the convolution and returning maximum response as a return.

@@

\subsection{Object Detection} 
Instance-level 2D-3D matching (registration) is a difficult problem since we need a CAD model that looks similar to the object in the image. Especially, due to extreme intraclass variability of chair \cite{Aubry14} used a very large collection of CAD models to cover variety of chairs. Similarly, \cite{Lim14}
Our ensemble of NZ-WHO templates is generated from rendering image with no calibration but shows strong performance on object detection. The baseline method are Aspect Layout Model (ALM) \cite{Xiang12} and VOC-DPM \cite{Pepik12}. Note that the baseline methods require complex 3D modeling and viewpoint annotated natural images.

@@

@@

\begin{table*}[!htbp]
    \footnotesize
  \begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
AP/AVP              & V-DPM \cite{Xiang14} & VOC-DPM \cite{Pepik12}  & \cite{Pepik12} + Ours (discrete) & \cite{Pepik12} + Ours (full) & R-CNN + Ours (discrete) & R-CNN + Ours (full)\\
\hline\hline
car-4v              & 37.2 / 20.2         & 45.6 / 36.9             & 47.6 / 42.7       & 47.6 / \textbf{42.7}        & 49.6 / 41.5         & 49.6 / 41.5\\ \hline
car-8v              & 37.3 / 23.5         & 47.6 / 36.6             & 47.6 / \textbf{39.8}       & 47.6 / 39.5        & 49.6 / 38.0         & 49.6 / 39.0\\ \hline
car-16v             & 36.6 / 18.1         & 46.0 / 29.6             & 47.6 / 32.7       & 47.6 / 33.0        & 49.6 / 34.0         & 49.6 / \textbf{34.3}\\ \hline
car-24v             & 36.3 / 13.7         & 42.1 / 24.6             & 47.6 / 27.4       & 47.6 / 27.4        & 49.6 / 27.0         & 49.6 / \textbf{27.6}\\ \hline
\hline
bicycle-4v          & 45.2 / 41.7         & 46.9 / 43.9             & 48.1 / 47.6       & 48.1 / 46.6        & 61.7 / 56.5         & 61.7 / \textbf{56.7}\\ \hline
bicycle-8v          & 47.3 / 36.7         & 48.1 / 40.3             & 48.1 / 40.6       & 48.1 / 40.6        & 61.7 / 48.9         & 61.7 / \textbf{49.2}\\ \hline
bicycle-16v         & 46.5 / 18.4         & 45.6 / 22.9             & 48.1 / 26.2       & 40.1 / 27.3        & 61.7 / 34.7         & 61.7 / \textbf{35.8}\\ \hline
bicycle-24v         & 44.4 / 14.3         & 45.9 / 16.7             & 48.1 / 21.5       & 40.1 / 20.9        & 61.7 / \textbf{27.0}         & 61.7 / 23.9\\ \hline
\end{tabular}
\end{center}
\caption{Average Precision(AP) and Average Viewpoint Precision(AVP) on PASCAL 3D dataset\cite{Xiang14}. For combined method (* + Ours), we use bounding boxes from * and augment viewpoint using our method. R-CNN and run our method to produce 2D-3D matching. If our method fails to give viewpoint, it predicts the viewpoint to be 0. Note that our method gives high quality metadata such as continuous 3D viewpoint, CAD model (rendering depth) and fine-grained category, whereas base-line methods gives 1D discrete azimuths. Our R-CNN was not an optimized version so the detection AP is lower than the state-of-the-art R-CNN performance}
\label{tab:pascal12}
\end{table*}

>>>>

%   In contrast to other mid-level patch based 3D matching methods which requires extensive training and more than 10Gb of memory space and physical storage, ours require no training nor saving the templates thus enabling it to run on any personal computer. 
%   To overcome these difficulties, we propose an efficient and fast way to synthesize and validate templates using on the fly realistic rendering that can jointly optimize scale, translation, rotation and focal length continuously. We render an object and extract features and decorrelate spatial dependencies within them on the fly to make a discriminative template. To speed up the decorrelation procedure, we adopt a gradient based algorithm which is numerically more stable than decomposition procedure. Finally, we efficiently search for candidate match by computing convolution using FFT and jointly optimize to match the object accurately. To speed up further, we propose an efficient GPU implementation and tested on PASCAL images.

\end{abstract}

@@

\textcolor{red}{Should we mention how our method improves on this (ie by using CG)?}

% Using HOG
% In this paper, we also used Histograms of Orientations feature \cite{Dalal05} but in general, we could use any dense features.

@@

\begin{table*}[!htbp]
    \footnotesize
    \begin{center}
\begin{tabular}{|c|c|r|c|r|}
\hline
Methods (AP/MPPE) & before calibration  & time & after calibration \cite{Aubry14} & time \\
\hline\hline
HOG\cite{Dalal05}     & 72.3 / 65.0           &  31ms  & 60.4 / 50.2                 & 8.7 sec \\ 
WHO\cite{Hariharan12} & 82.1 / 85.4           &  3811ms& 84.4 / 83.0                 & 12.4 sec  \\
WHO-CG                & 81.7 / \textbf{84.9}  &  104ms & 83.7 / 87.3                 & 8.3 sec \\
WHO-CG-Z              & 54.4 / 65.1           &  103ms & \textbf{92.8} / 86.7        & 8.7 sec  \\
% NZC-WHO-Z    & 89.10/\textbf{78.64} &    & 91.15/74.79                  &     \\
NZ-WHO-CG             & \textbf{90.0} / 82.8  &   79ms & 90.3 / \textbf{86.8}        & 8.5 sec   \\
\hline
\end{tabular}
\end{center}
\caption{Average Precision(AP) and Mean Precision in Pose Estimation (MPPE) \cite{Lopez-Sastre11} variations of WHO on 3DObject Car dataset\cite{Savarese07}. WHO refers to standard WHO using the method presented in \cite{Hariharan12}, WHO-CG uses iterative Conjugate Gradient method to generate WHO. WHO-CG-Z uses whiten the whole template and zero out textureless region. NZ-WHO-CG is the NZ-WHO which whitens only non-zero cells using iterative Conjugate Gradient method. The time column indicates the time to generate one template. We followed calibration procedure presented in \cite{Aubry14}.}
\label{tab:who_initializations}
\end{table*}

@@

% Estimating viewpoint continuously to fine tune a initial matching
The NZ-WHO template matching method we have presented makes template generation and evaluation computationally inexpensive. 
Since creating a template and validating the template become computationally inexpensive, making a proposal and validating the proposal to fine tune a 2D-3D matching became affordable. We used Metropolis-Hastings algorithm to generate a proposal pose and focal length and validate the proposal using the convolution and returning maximum response as a return.

@@

Second, we show that our method can complement the detection bounding boxes coming out of standard object detectors. This process will enrich object detection bounding boxes with 2D-3D matching which gives continuous viewpoint estimation, a CAD model (rendering, depth) and fine-grained category.

\subsection{2D-3D Matching as an Object Detector}

We created ensemble of NZ-WHO templates using 9 CAD models and measured the Average Precision (AP) of object detection task and Mean Precision in Pose Estimation (MPPE) \cite{Lopez-Sastre11}. MPPE is the average of viewpoint accuracy.

% Instance-level 2D-3D matching (registration) is a difficult problem since we need a CAD model that looks similar to the object in the image. Especially, due to extreme intraclass variability of chair \cite{Aubry14} used a very large collection of CAD models to cover variety of chairs. Similarly, \cite{Lim14}

Our ensemble of NZ-WHO templates is generated from rendering images within few minutes but shows reasonable performance on object detection task. The baseline method are Aspect Layout Model (ALM) \cite{Xiang12} and VOC-DPM \cite{Pepik12}. Note that the baseline methods require complex 3D modeling and viewpoint annotated natural images.

@@

\begin{table}[!htbp]
    \footnotesize
  \begin{center}
    \begin{tabular}{|c|c|c|c|}
    \hline
     AP/MPPE& Ours & ALM\cite{Xiang12} & VOC-DPM\cite{Pepik12} \\
    \hline\hline
    car & \textbf{99.8}/91.7 &  98.4/93.4 & \textbf{99.8}/\textbf{97.5} \\ 
    bicycle & 93.0/87.4 & 93.0/91.4 & \textbf{98.8}/\textbf{97.5} \\
    \hline
    \end{tabular}
  \end{center}
  \caption{Average Precision(AP) and Mean Precision in Pose Estimation (MPPE) on 3DObject dataset car category. Our method produces high quality 2D-3D matching yet performs on par with state-of-the art detectors.}
  \label{tab:3dobject}
\end{table}

@@

We used detection bounding boxes and augment it using our method to give viewpoint estimation. We used detection bounding boxes from two different methods: VOC-DPM \cite{Pepik12} and R-CNN \cite{Girshick14}. VOC-DPM is a variation of DPM that can estimate viewpoint and is the state-of-the-art performance in viewpoint estimation \cite{Xiang14}. The result is on Table \ref{tab:pascal12} and precision recall curves for car using R-CNN + Ours are on Figure \ref{fig:car_cnn_ap}.

We got significant improvement on viewpoint estimation by using our method on top of general purpose detectors. Note that our method estimates azimuth, elevation, in-plane rotation, focal length estimation, fine-grained categorization whereas V-DPM, VOC-DPM only estimate discrete viewpoints.

Example outputs are on Figure \ref{fig:pascal12cnn}.

\begin{figure*}[h]
\setlength\tabcolsep{1pt}
\centering
\begin{tabular}{|c|c|c|c|c|}
   \hline
  \includegraphics[width=0.24\textwidth]{car_cnn/1a.png} &   
  \includegraphics[width=0.24\textwidth]{car_cnn/1b.png} &   
  \includegraphics[width=0.24\textwidth]{car_cnn/1c.png} &   
  \includegraphics[width=0.24\textwidth]{car_cnn/1d.png}  \\  
   \hline
  \includegraphics[width=0.24\textwidth]{car_cnn/2a.png} &   
  \includegraphics[width=0.24\textwidth]{car_cnn/2b.png} &   
  \includegraphics[width=0.24\textwidth]{car_cnn/2c.png} &   
  \includegraphics[width=0.24\textwidth]{car_cnn/2e.png}  \\  
   \hline
  \includegraphics[width=0.24\textwidth]{car_cnn/4a.png} &   
  \includegraphics[width=0.24\textwidth]{car_cnn/4b.png} &   
  \includegraphics[width=0.24\textwidth]{car_cnn/4c.png} &   
  \includegraphics[width=0.24\textwidth]{car_cnn/4d.png}  \\  
%   \includegraphics[width=0.24\textwidth]{car_cnn/10a.png} &   
%   \includegraphics[width=0.24\textwidth]{car_cnn/10b.png} &   
%   \includegraphics[width=0.24\textwidth]{car_cnn/10c.png} &   
%   \includegraphics[width=0.24\textwidth]{car_cnn/10d.png}  \\  
%   \includegraphics[width=0.24\textwidth]{bicycle_cnn/3a.png} &   
%   \includegraphics[width=0.24\textwidth]{bicycle_cnn/3b.png} &   
%   \includegraphics[width=0.24\textwidth]{bicycle_cnn/3c.png} &   
%   \includegraphics[width=0.24\textwidth]{bicycle_cnn/3d.png}  \\
  \hline
  \includegraphics[width=0.24\textwidth]{bicycle_cnn/7a.png} &   
  \includegraphics[width=0.24\textwidth]{bicycle_cnn/7b.png} &   
  &\\
  \hline
  \includegraphics[width=0.24\textwidth]{bicycle_cnn/4a.png} &   
  \includegraphics[width=0.24\textwidth]{bicycle_cnn/4b.png} &   
  \includegraphics[width=0.24\textwidth]{bicycle_cnn/4c.png} &   
  \includegraphics[width=0.24\textwidth]{bicycle_cnn/4d.png}  \\
  \hline
  CNN Proposals & \multicolumn{3}{|c|}{Our matching results on proposal bounding boxes} \\
  \hline
\end{tabular}
\caption{Example enriched bounding boxes. Given R-CNN\cite{Girshick14} detection bounding boxes, our method predicted 2D-3D matching reasonably. On the first column, R-CNN detection bounding boxes overlaid. From the second to the last columns, output of our method given a R-CNN box. Blue boxes are R-CNN output and purple boxes are the tightest bounding box enclosing predicted CAD model}
  \label{fig:pascal12cnn}
\end{figure*}
\begin{table*}[!htbp]
    \footnotesize
  \begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
AP/AVP              & V-DPM \cite{Xiang14} & VOC-DPM \cite{Pepik12}  & \cite{Pepik12} + Ours (discrete) & \cite{Pepik12} + Ours (full) & R-CNN + Ours (discrete) & R-CNN + Ours (full)\\
\hline\hline
car-4v              & 37.2 / 20.2         & 45.6 / 36.9             & 47.6 / 42.7       & 47.6 / \textbf{42.7}        & 49.6 / 41.5         & 49.6 / 41.5\\ \hline
car-8v              & 37.3 / 23.5         & 47.6 / 36.6             & 47.6 / \textbf{39.8}       & 47.6 / 39.5        & 49.6 / 38.0         & 49.6 / 39.0\\ \hline
car-16v             & 36.6 / 18.1         & 46.0 / 29.6             & 47.6 / 32.7       & 47.6 / 33.0        & 49.6 / 34.0         & 49.6 / \textbf{34.3}\\ \hline
car-24v             & 36.3 / 13.7         & 42.1 / 24.6             & 47.6 / 27.4       & 47.6 / 27.4        & 49.6 / 27.0         & 49.6 / \textbf{27.6}\\ \hline
\hline
bicycle-4v          & 45.2 / 41.7         & 46.9 / 43.9             & 48.1 / 47.6       & 48.1 / 46.6        & 61.7 / 56.5         & 61.7 / \textbf{56.7}\\ \hline
bicycle-8v          & 47.3 / 36.7         & 48.1 / 40.3             & 48.1 / 40.6       & 48.1 / 40.6        & 61.7 / 48.9         & 61.7 / \textbf{49.2}\\ \hline
bicycle-16v         & 46.5 / 18.4         & 45.6 / 22.9             & 48.1 / 26.2       & 40.1 / 27.3        & 61.7 / 34.7         & 61.7 / \textbf{35.8}\\ \hline
bicycle-24v         & 44.4 / 14.3         & 45.9 / 16.7             & 48.1 / 21.5       & 40.1 / 20.9        & 61.7 / \textbf{27.0}         & 61.7 / 23.9\\ \hline
\end{tabular}
\end{center}
\caption{Average Precision(AP) and Average Viewpoint Precision(AVP) on PASCAL 3D dataset\cite{Xiang14}. For combined method (* + Ours), we use bounding boxes from * and augment viewpoint using our method. R-CNN and run our method to produce 2D-3D matching. If our method fails to give viewpoint, it predicts the viewpoint to be 0. Note that our method gives high quality metadata such as continuous 3D viewpoint, CAD model (rendering depth) and fine-grained category, whereas base-line methods gives 1D discrete azimuths. Our R-CNN was not an optimized version so the detection AP is lower than the state-of-the-art R-CNN performance}
\label{tab:pascal12}
\end{table*}

\begin{figure}[h]
\setlength\tabcolsep{1pt}
\centering
\begin{tabular}{cc}
  \hline
  \includegraphics[width=0.49\linewidth]{car_cnn4_crop.png} &   
  \includegraphics[width=0.49\linewidth]{car_cnn8_crop.png} \\   
  \includegraphics[width=0.49\linewidth]{car_cnn16_crop.png} &   
  \includegraphics[width=0.49\linewidth]{car_cnn24_crop.png} \\   
  \hline
\end{tabular}
\caption{Average Precision (AP) and Average Viewpoint Precision (AVP) and viewpoint confusion table on PASCAL 12 car validation set using R-CNN + Ours. From the top left, 4 views, 8 views, 16 views and 24 views. Since our method only augment the object detection, the AP remains the same for all cases. As we increase the viewpoint estimation, the AVP decreases. Note that since we estimate null viewpoint if our method fails to find match, there are more 0 viewpoint predictions (top row of confusion matrices) }
  \label{fig:car_cnn_ap}
\end{figure}
